Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-04-02 thesis/chapters/ch_chapter4.tex
Finished text (mostly) to chapter 4 with exception of the introduction. Just need to re-read and regenerate plots with updated results

@@

\subsection{Mixing Experiment Setup}
\label{ch4:sec:experiment}
We restart our fiducial, full-physics simulation at three different times, 100 Myr, 180 Myr, and 360 Myr; runs at these times are labelled as I, II, and III respectively. These correspond to three different times in the galaxy's SFR evolution, testing how much variance is expected in the metal mixing and ejection with the star formation rate. I occurs in the last $\sim$ 20 Myr of the initial burst of star formation, II occur during the lull in star formation following this peak, and III in an extended period of little to no ongoing star formation. We attempted to to evolve each simulation for 150 Myr, but due to computational constrains (particularly with the I runs) this was not always possible.

At the beginning of each restart, we place by-hand one or more enrichment events at assigned positions throughout the galaxy, with thermal injection energies ($E_{\rm ej}$), masses ($m_{\rm ej}$), and metal fractions ($Z_{\rm ej}$). We note that the injection masses, particularly the metal fractions, are somewhat arbitrary as they are likely dynamically insignificant relative to the ambient ISM mass in which they occur. The important parameter here is $E_{\rm ej}$, which we vary to sample the range of ejection energies associated with significant sources of chemical enrichment, including AGB winds ($10^{46}$~erg)\footnote{Assuming full thermalization of the total mechanical energy output of an AGB wind.}, NS-NS mergers ($10^{49} - 10^{50}$~erg), supernovae (10$^{51}$~erg), and exotic enrichment sources, such as hypernovae, that can reach much higher energies (10$^{52}$~erg). Each run contains only sources from a single event type, as indicated in the run-name by the log of the injection energy in ergs. For example, the run beginning at 180~Myr with AGB-like events is labeled ``II\_E46". The metal enrichment from each source is tracked and evolved with a unique passive scalar tied to the individual source and separate from any additional chemical enrichment that may occur within the galaxy over time.

@@

We focus this analysis primarily on the results from run II, and save a discussion of how these results vary across runs and across specific source positions in Section XXX \textit{(maybe take out if don't have time)}.

@@

We summarize the source-averaged results for all enrichment events in runs II\_E46, II\_E49, II\_E50, II\_E51, and II\_E52 in Figure~\ref{ch4:fig:ISM_CGM} by showing the fraction of source metals contained within each phase of the ISM (colored lines, which sum to the black dashed line), and the CGM (black, solid). These plots show a clear, immediate trend across $E_{\rm ej}$ for all lines in the figure, but the differences are most striking for the ISM (black, solid), CGM (black, dashed) lines. In Chapter~\ref{ch:chapter3}
% \citep{Emerick2019}
we demonstrated that there was a significant difference in metal ejection fraction, $f_{\rm ej}$, for metals from AGB sources as compared to metals released in SNe. Here, we confirm that this is driven predominantly by differences in the energy of the events, and not necessarily when or where they occur in the ISM. The clear trend in Figure~\ref{ch4:fig:ISM_CGM} shows that events with higher $E_{\rm ej}$ are much more readily ejected from the disk of the galaxy, while events with lower $E_{\rm ej}$ couple poorly to the same galactic outflows.
\textit{statement on how immediate enrichment is after events}
% By 10~Myr after the enrichment events,
The high-energy events rapidly converge on a peak $f_{\rm ej}$ within about 20~Myr of the event, with only a gradual increase towards the end of the 150~Myr as the metals in the ISM are swept up in additional outflows. The lower energy events evolve more gradually. The lowest energy event, corresponding to AGB winds, reaches $f_{\rm ej}$ ~ 0.68, which increases to 0.87 for the 10$^{51}$~erg events, and 0.95 for the 10$^{52}$~erg event. To emphasize these differences, we plot each line separately in the left panel of Figure~\ref{ch4:fig:CGM_CNM}

This implies that, if one to were assume that all metals behaved the same and adopted an $f_{\rm ej}$ of $\sim$0.9 -- typical for the total metal ejection fraction in simulations that do not track individual enrichment channels \citep[e.g.][]{Muratov2017,Christensen2018}, and consistent with observations of the O abundance in Local Group dwarf galaxies \citep[e.g.][]{Kirby2011,McQuinn2015} -- one would underestimate the amount of AGB metals in the ISM by a factor of $\sim$ 3, and overestimate exotic, high-energy sources by a factor of $\sim$ 2. In each case, these are metals that cannot participate in the enrichment of future stellar populations.

In all cases, the metals contained with the ISM are deposited predominately in the ionized phases of the ISM, the WIM or the HIM, with the relative fraction in each phase driven by the energy of the event. Metals in the ISM injected with $E_{\rm ej} > 10^{49}$~erg are initially located predominately in the HIM, roughly 1/2 for $E_{\rm ej} = 10^{49}$~erg and nearly all for $E_{\rm ej} > 10^{52}$~erg. The lowest energy events are initially in the WIM and WNM, tracing the two dominant volume-filling components of the ISM (see Chapter~\ref{ch:chapter1}), as these events do not have sufficient energy to generate the HIM phase by themselves.

@@

What we can examine, however, is the final quantity of interest: the long-term evolution of the metals from each of these sources, and how long it takes for these metals to propagate from the warm / hot phases of the ISM to cold, star forming gas. In the right panel of Figure~\ref{ch4:fig:CGM_CNM}, we examine the evolution of the metal fraction of the CNM \textit{for just those elements retained in the ISM}. Although the initial CNM fractions are about the same for each source, the evolution over the first $\sim$50~Myr is qualitatively different. the higher energy sources, $E_{\rm ej} > 10^{49}$~erg, are more rapidly incorporated into the CNM than E46, even though each retains a lower fraction of metals in the ISM. This is most significant at $\sim$20~Myr, when the fraction of metals in the CNM from these sources is a factor of $\sim$3-4 higher than the E46 metals. By $\sim$50~Myr, the evolutions become intermingled and complex, with no clear trend as a function of $E_{\rm ej}$.

@@

However, the metals from each source do gradually trend towards a fraction of $\sim$0.8 by the end of the 150~Myr simulation time, corresponding to the mass fraction of the CNM at this point in the galaxy's evolution (see Figure~\ref{ch1:fig:}). This trend -- that the fraction of metals contained in a given phases tends towards the mass fraction of that phase -- is true across all phases in the simulation. Therefore, we can consider that the metals for each source are well-mixed across the \textit{phases} of the ISM on timescales of $\sim$100-150~Myr. \textbf{Compute the cooling timescale of each phase (maybe with a phase diagram?) and also the dynamical timescale and make comparisons here.}. These two results suggest that the delay-time between when metals are available to enrich ongoing star formation is on order of $\sim$10~Myr across sources, and only significant for the lowest (E46, AGB-like) energy sources. Although well-mixed across phases, we emphasize that this does not imply that the abundances are the same across phases nor that the metals are spatially well-mixed across the galaxy. We investigate this second point further below.

@@

To frame this analysis, lets first examine what processes affect the metal mass fraction PDF. The mean metal mass fraction for a collection of gas is simply the total metal mass divided by the total gas mass. For a collection of gas elements, the mean metal mass fraction is defined as $Z_{\rm gas} = \frac{ \Sigma Z_{\rm i}M_{\rm gas,i}}{\Sigma M_{\rm gas,i}}$, where $M_{\rm gas,i}$ is the mass of the i$^{\rm th}$ homogeneous blob of gas, and $Z_{i}$ its metal mass fraction. A completely homogeneous ISM would have a $\delta$ function metal mass fraction PDF at the mean metal fraction, or $p(Z) = \delta(Z_{\rm gas})$. By definition, then, a homogeneous distribution will have a zero-width IQR, IDR, and mean-median difference. Non-uniform enrichment in a homogeneous medium will increase each of these quantities, while mixing in a non-uniform medium will tend to decrease each. In simple models of galactic chemical evolution, the mean gas abundance -- not the median -- is followed, even though the median is more indicative of the abundance of a ``typical" gas element. Because these two quantities provide a more immediate conceptual understanding of how metals are distributed in the ISM, we focus on characterizing the inhomogeneity of the followed metals by examining the mean-median difference. Though this quantity being zero is necessary for metals to be homogeneously distributed in a medium, we do note that it is not a sufficient condition. However, as noted, its evolution is easy to interpret conceptually. Enrichment increases this quantity by raising the mean metal fraction while keeping the median value fixed (or only slightly increasing), so long as the newly enriched mass is a small fraction of the total gas mass (as is usually the case). Conversely, preferential removal of metals from a medium will lower this quantity. And finally, the mixing of metal-rich gas elements with metal-poor will gradually bring the mean and median values to parity. For our numerical experiments, enrichment in the whole ISM only occurs once, while ejection from the ISM is easy to follow. This allows for a more clean interpretation of the evolution than examining either the IQR or the IDR.

\begin{figure*}
  \centering
  \includegraphics[width=0.75\linewidth]{figures/ch4/ISM_CNM_average_mean-median}
  \caption{Time evolution of the difference (in dex) between the mean metal mass fraction of each source and the median metal mass fraction of each source in all of the gas in the ISM (left) and in the CNM only (right).}
  \label{ch4:fig:CGM_CNM}
\end{figure*}

In Figure~\ref{ch4:fig:mean-median} we examine the average behavior of the mean-median metal mass fractions for the whole ISM in each run (left) and the CNM only (right). In all cases, enrichment drives the dramatic increase of this value from zero to over 10 dex. This large value is due entirely because the initial abundance for each metal is zero. This is followed by a significant decrease in spread over he first 25~Myr or so as metals are preferentially ejected from the ISM via outflows. The delay in this spike in the CNM (compared to the galaxy as a whole) is due to the different timescales over which most of the metals are incorporated into the cold gas, as shown in Figure~\ref{ch4:fig:ISM_CGM}.

Clearly, at no point can any of these distributions be considered well-mixed within the 150~Myr simulation time. \textit{Need to fix up: This is in agreement with the understanding that the mxiing timescales should correspond to XXX, which suggests that it would take $\sim$X~Gyr (1 Gyr?) for the metals to be fully mixed}. Interestingly, however, the injection energy for a given sources leaves an indelible impact on the subsequent evolution of those metals in the ISM throughout the simulation time. We demonstrate that variations in abundance spread across elements may be driven significantly by the enrichment site and its associated $E_{\rm ej}$. Metals from II\_E46 are significantly less well-mixed than their higher-energy counterparts at II\_E51 and II\_E52. Interestingly, II\_E49 behaves much more similarly to II\_E46 -- with poor mixing -- even though it is much closer in energy to the II\_E51. For what mixing does occur in the CNM for these elements, it is apparent that it happens much faster ($\sim$20~Myr) for the highest energy event than the lower energy events which take longer to plateau to a consistent value.

We note that these results are given for \textit{single} enrichment events. If many events occur throughout the galaxy with similar metal yields then this would tend to decrease the abundance variations across the galaxy. However, the AGB wind metals examined in Chapter~\ref{ch:chapter3} are still more heterogeneously distributed than core collapse SNe elements in spite of the fact that these events are both more common than SNe and are more evenly distributed throughout the galaxy.

@@

\subsection{Event-by-Event Variation}

For clarity we focused on the \textit{mean} evolution of individual enrichment events at fixed injection energy in the results thus far, but we turn now to discuss how much variety exists among these individual events. For each of the discussed runs (from set II), Figure~\ref{ch4:fig:CGM_CNM_variance} shows the fraction of metals ejected from the galaxy (left) and the fraction of metals in the ISM within the CNM (center) as a function of source energy at two different times, 20~Myr (blue) and 100~Myr (orange). In the right panel, we show how the spread at each of these times varies for each event. The averaged values at each energy are shown as points, and the bars denote the min/max value for each event energy.\footnote{While the standard deviation or IQR may be more informative here, rather than min / max, the sample size at each energy, with the exception of the E46 runs, is too small to reliably do so.} As shown, there is substantial variation at each energy. This is particularly striking for the E46 events, which seemingly can have effectively any value for $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ at any point during the evolution. Finally, the right panel shows that the degree of inhomogeneous mixing for each event varies significantly as well. However, there is a more clear trend for each of the mean, min, and max values of the spread at 100~Myr (decreasing as a function of increasing $E_{\rm ej}$).

 \begin{figure*}
   \centering
   \includegraphics[width=0.8\linewidth]{figures/ch4/II_Eej_CNM_avg}
   \caption{A summary of the results presented in Figure~\ref{ch4:fig:CGM_CNM} and Figure~\ref{ch4:fig:CGM_CNM} that demonstrates the significant variation in $f_{\rm CGM}$, $f_{\rm CNM} / f_{\rm Disk}$, and the mean-median separate at fixed injection energy. The points give the average values at each energy at two different times, while the error bars show the min/max values at each energy. The points at different times have been offset slightly along the horizontal axes for clarity. Since they are only two points, we show the results from II\_E52\_r0 and II\_E52\_r300 without averaging.}
   \label{ch4:fig:CGM_CNM_variance}
 \end{figure*}

@@

\subsection{Dependence on Radial Position}
\label{ch4:sec:radial position}
We place each source in regular, but arbitrary, positions in the galaxy without consideration for the local ISM conditions at injection. Sources are placed at the center of the galaxy ($r = 0$~pc) or at various radii (at cardinal positions at 300~pc, and 600~pc), all in the mid-place of the galaxy. Since the E46 events were likely to not self-interact, we additionally placed events at $r = 100$~pc and one event at each $r$ at about one scale height above the disk ($z = 50$~pc). For each II run, we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ as a function of radial position in the galaxy for each event in Figure~\ref{ch4:fig:radial position}. The points, error bars, and colors are the same as in Figure~\ref{ch4:fig:CGM_CNM_variance}.

@@

We compute the average ISM properties within a 4-zone radius ($4 \times dx$=7.2~pc) around each injection site to examine any potential correlations in evolution with ISM properties.\footnote{Note that the injection region for each event is a 3-cell radius sphere mapped onto the grid with a CIC interpolation scheme.} In Figure~\ref{ch4:fig:ISM_variance} we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ for each II run as a function of the average number density ($<n>$) in the injection region within 0.1~Myr (our time resolution) of the event. The E46 events clearly do a better job sampling the ISM density PDF, with a slight trend of decreasing $f_{\rm CGM}$ (more retained metals) with local ISM number density. This difference is even true after 100~Myr of evoltuion, but still with significant variation at fixed $n$. The number statistics limit a conclusive result for the higher energy events, but generally there does not appear to be an obvious trend in behavior of $f_{\rm CGM}$ as a function of $n$. This is perhaps because these energies are sufficient to heat up and dramatically reduce the density of the ISM in the injection region anyway, such that $n$ before the event is not as important as it is for the E46 events which are much less dynamically important. $f_{\rm CNM} / f_{\rm Disk}$ does not display any significant variation with $n$ at 100~Myr, but is slightly elevated at 20~Myr at higher densities. There is still a variety of behavior at fixed $n$. Finally, we note no obvious trends of the spread of the metal PDFs as a function of $n$ for all but the E46 run. Again, the local ISM densities is likely not too dynamically relevant to the higher energy events -- hence their lack of variation with $n$ -- but this is not true for E46. There is a fairly clear trend of increasing inhomogeneity in the CNM as a function of the initial $n$ in the injection region for E46, that is not present at 20~Myr, but is obvious at 100~Myr.

@@

Finally, we examine the companion runs (I and III) to see if the global star formation rate of the galaxy drives variation in these results. Although this does not help to directly account for the variation seen in the previous plots, this could provide insight into how SFR (in an averaged sense) affects the evolution of metals. We note that correlation with SFR is of interest not because the formation of stars by themselves, but rather the increase in feedback associated with a higher SFR (and thus a warmer / hotter ISM, greater turbulence in the ISM, and more significant outflows).

@@

Galactic chemical evolution is far from ``one-zone" and metal mixing is clearly neither homogeneous nor instantaneous. In addition, the

>>>>

\newcommand{\runone}{I\_}
\newcommand{\runonenu}{I}
\newcommand{\runtwo}{II\_}
\newcommand{\runtwonu}{II}

@@

\subsection{Mixing Experiment Setup}
\label{ch4:sec:experiment}
We restart our fiducial, full-physics simulation at two different times, 180 Myr, and 360 Myr, labeled as runs \runonenu~ and \runtwonu~ respectively. These correspond to two different times in the galaxy's SFR evolution, testing how much variance is expected in the metal mixing and ejection with the star formation rate. \runonenu~ occurs during the lull in star formation following the initial SFR peak, and \runtwonu~ occurs in an extended period of little to no ongoing star formation. We attempted to to evolve each simulation for 150 Myr, but due to computational constrains this was not always possible.\footnote{We began runs during the peak of star formation ($\sim 100$~Myr), but the computational cost increases substantially during this time. These runs were abandoned.}

% I occurs in the last $\sim$ 20 Myr of the initial burst of star formation,

At the beginning of each restart, we place by-hand one or more enrichment events at assigned positions throughout the galaxy, with thermal injection energies ($E_{\rm ej}$), masses ($m_{\rm ej}$), and metal fractions ($Z_{\rm ej}$). We note that the injection masses, particularly the metal fractions, are somewhat arbitrary as they are likely dynamically insignificant relative to the ambient ISM mass in which they occur. The important parameter here is $E_{\rm ej}$, which we vary to sample the range of ejection energies associated with significant sources of chemical enrichment, including AGB winds ($10^{46}$~erg)\footnote{Assuming full thermalization of the total mechanical energy output of an AGB wind.}, NS-NS mergers ($10^{49} - 10^{50}$~erg), supernovae (10$^{51}$~erg), and exotic enrichment sources, such as hypernovae, that can reach much higher energies (10$^{52}$~erg). Each run contains only sources from a single event type, as indicated in the run-name by the log of the injection energy in ergs. For example, the run beginning at 180~Myr with AGB-like events is labeled ``\runone E46". The metal enrichment from each source is tracked and evolved with a unique passive scalar tied to the individual source and separate from any additional chemical enrichment that may occur within the galaxy over time.

@@

@@

We summarize the source-averaged results for all enrichment events in runs \runone E46, \runone E49, \runone E50, \runone E51, and \runone E52 in Figure~\ref{ch4:fig:ISM_CGM} by showing the fraction of source metals contained within each phase of the ISM (colored lines, which sum to the black dashed line), and the CGM (black, solid). These plots show a clear, immediate trend across $E_{\rm ej}$ for all lines in the figure, but the differences are most striking for the ISM (black, solid), CGM (black, dashed) lines. In Chapter~\ref{ch:chapter3}
% \citep{Emerick2019}
we demonstrated that there was a significant difference in metal ejection fraction, $f_{\rm ej}$, for metals from AGB sources as compared to metals released in SNe. Here, we confirm that this is driven predominantly by differences in the energy of the events, and not necessarily when or where they occur in the ISM. The clear trend in Figure~\ref{ch4:fig:ISM_CGM} shows that events with higher $E_{\rm ej}$ are much more readily ejected from the disk of the galaxy, while events with lower $E_{\rm ej}$ couple poorly to the same galactic outflows.
%\textit{statement on how immediate enrichment is after events}
% By 10~Myr after the enrichment events,
The high-energy events rapidly converge on a peak $f_{\rm ej}$ within about 20~Myr of the event, with only a gradual increase towards the end of the 150~Myr as the metals in the ISM are swept up in additional outflows. The lower energy events evolve more gradually. The lowest energy event, corresponding to AGB winds, reaches $f_{\rm ej}$ ~ 0.68, which increases to 0.87 for the 10$^{51}$~erg events, and 0.95 for the 10$^{52}$~erg event. To emphasize these differences, we plot each line separately in the left panel of Figure~\ref{ch4:fig:CGM_CNM}

This implies that, if one to were assume that all metals behaved the same and adopted an $f_{\rm ej}$ of $\sim$0.9 -- typical for the total metal ejection fraction in simulations that do not track individual enrichment channels \citep[e.g.][]{Muratov2017,Christensen2018}, and consistent with observations of the O abundance in Local Group dwarf galaxies \citep[e.g.][]{Kirby2011,McQuinn2015} -- one would underestimate the amount of AGB metals in the ISM by a factor of $\sim$ 3, and overestimate exotic, high-energy sources by a factor of $\sim$ 2. In each case, these are metals that cannot participate in the enrichment of future stellar populations and will make a significant difference on the output stellar abundances in chemical evolution models. We explore this impact in different works.

In all cases, the metals contained with the ISM are deposited predominately in the ionized phases of the ISM, the WIM or the HIM, with the relative fraction in each phase driven by the energy of the event. Metals in the ISM injected with $E_{\rm ej} > 10^{49}$~erg are initially located predominately in the HIM, roughly 1/2 for $E_{\rm ej} = 10^{49}$~erg and nearly all for $E_{\rm ej} > 10^{51}$~erg. The lowest energy events are initially in the WIM and WNM, tracing the two dominant volume-filling components of the ISM (see Chapter~\ref{ch:chapter1}), as these events do not have sufficient energy to generate the HIM by themselves.

@@

What we can examine, however, is the final quantity of interest: the long-term evolution of the metals from each of these sources, and how long it takes for these metals to propagate from the warm / hot phases of the ISM to cold, star forming gas. In the right panel of Figure~\ref{ch4:fig:CGM_CNM}, we examine the evolution of the metal fraction of the CNM \textit{for just those elements retained in the ISM}. Although the initial CNM fractions are about the same for each source, the evolution over the first $\sim$50~Myr is qualitatively different. The higher energy sources, $E_{\rm ej} > 10^{49}$~erg, are more rapidly incorporated into the CNM than E46, even though each retains a lower fraction of metals in the ISM. This is most significant at $\sim$20~Myr, when the fraction of metals in the CNM from these sources is a factor of $\sim$3-4 higher than the E46 metals. By $\sim$50~Myr, the evolutions become intermingled and complex, with no clear trend as a function of $E_{\rm ej}$.

@@

However, the metals from each source do gradually trend towards a fraction of $\sim$0.8 by the end of the 150~Myr simulation time, corresponding to the mass fraction of the CNM at this point in the galaxy's evolution (see Figure~\ref{ch1:figISM_evolution}). This trend -- that the fraction of metals contained in a given phases tends towards the mass fraction of that phase -- is true across all phases in the simulation. Therefore, we can consider that the metals for each source are well-mixed across the \textit{phases} of the ISM on timescales of $\sim$100-150~Myr. \textbf{Compute the cooling timescale of each phase (maybe with a phase diagram?) and also the dynamical timescale and make comparisons here.}. These two results suggest that the delay-time between when metals are available to enrich ongoing star formation is on order of $\sim$10~Myr across sources, and only significant for the lowest (E46, AGB-like) energy sources. Although well-mixed across phases, we emphasize that this does not imply that the abundances are the same across phases nor that the metals are spatially well-mixed across the galaxy. We investigate this second point further below.

@@

To frame this analysis, lets first examine what processes affect the metal mass fraction PDF. The mean metal mass fraction for a collection of gas is simply the total metal mass divided by the total gas mass. For a collection of gas elements, the mean metal mass fraction is defined as $Z_{\rm gas} = \frac{ \Sigma Z_{\rm i}M_{\rm gas,i}}{\Sigma M_{\rm gas,i}}$, where $M_{\rm gas,i}$ is the mass of the i$^{\rm th}$ homogeneous blob of gas, and $Z_{i}$ its metal mass fraction. A completely homogeneous ISM would have a $\delta$ function metal mass fraction PDF at the mean metal fraction, or $p(Z) = \delta(Z_{\rm gas})$. By definition, then, a homogeneous distribution will have a zero-width IQR, IDR, and mean-median difference. Non-uniform enrichment in a homogeneous medium will increase each of these quantities, while mixing in a non-uniform medium will tend to decrease each. In simple models of galactic chemical evolution, the mean gas abundance -- not the median -- is followed, even though the median is more indicative of the abundance of a ``typical" gas element. Because these two quantities provide a more immediate conceptual understanding of how metals are distributed in the ISM, we focus on characterizing the inhomogeneity of the followed metals by examining the mean-median difference. Though it is necessary for this quantity to be zero for metals to be homogeneously distributed in a medium, we do note that it is not a sufficient condition (the mean of a symmetric distribution is equal to the median). However, its evolution is easy to interpret conceptually. Enrichment increases this quantity by raising the mean metal fraction while keeping the median value fixed (or only slightly increasing), so long as the newly enriched mass is a small fraction of the total gas mass (as is almost always the case). Conversely, preferential removal of metals from a medium will lower this quantity. And finally, the mixing of metal-rich gas elements with metal-poor will gradually bring the mean and median values to parity. For our numerical experiments, enrichment in the whole ISM only occurs once and ejection from the ISM is easy to follow. This allows for a more clean interpretation of the evolution than examining either the IQR or the IDR.

\begin{figure*}
  \centering
  \includegraphics[width=0.975\linewidth]{figures/ch4/ISM_CNM_average_mean-median}
  \caption{Time evolution of the difference (in dex) between the mean metal mass fraction of each source and the median metal mass fraction of each source in all of the gas in the ISM (left) and in the CNM only (right).}
  \label{ch4:fig:mean-median}
\end{figure*}

In Figure~\ref{ch4:fig:mean-median} we examine the average behavior of the mean-median metal mass fractions for the whole ISM in each run (left) and the CNM only (right). In all cases, enrichment drives the dramatic increase of this value from zero to over 10 dex. This large value is due entirely because the initial abundance for each metal is near zero and thus a majority of the gas (by mass) is at an arbitrarily small metal mass fraction. This is followed by a significant decrease in spread over the first 25~Myr or so as metals are preferentially ejected from the ISM via outflows. The delay in this spike in the CNM (compared to the galaxy as a whole) is due to the different timescales over which most of the metals are incorporated into the cold gas, as shown in Figure~\ref{ch4:fig:ISM_CGM}.

Clearly, at no point can any of these distributions be considered well-mixed within the 150~Myr simulation time. \textit{Need to fix up: This is in agreement with the understanding that the mxiing timescales should correspond to XXX, which suggests that it would take $\sim$X~Gyr (1 Gyr?) for the metals to be fully mixed}. Interestingly, however, the injection energy for a given source leaves an indelible impact on the subsequent evolution of those metals in the ISM throughout the simulation time. We demonstrate that variations in abundance spread across elements may be driven significantly by the enrichment site and its associated $E_{\rm ej}$. Metals from \runone E46 are significantly less well-mixed than their higher-energy counterparts at \runone E51 and \runone E52. Interestingly, \runone E49 behaves much more similarly to \runone E46 -- with poor mixing -- even though it is much closer in energy to the \runone E51. For what mixing does occur in the CNM for these elements, it is apparent that it happens much faster ($\sim$20~Myr) for the highest energy event than the lower energy events which take longer to plateau to a consistent value.

We note that these results are given for the average behavior of \textit{single} enrichment events. If many events occur throughout the galaxy with similar metal yields then this would tend to decrease the abundance variations across the galaxy. Indeed, the recovered metal mass fraction spreads from our fiducial simulations in Chapter~\ref{ch:chapter3} are much lower ($\sim$~1- 2 dex) than the $\sim$~2-5 dex widths we find here. However, it is still true that the the AGB wind metals examined in Chapter~\ref{ch:chapter3} are more heterogeneously distributed than core collapse SNe elements in spite of the fact that these events are both more common than SNe and are more evenly distributed throughout the galaxy.

@@

\subsection{Event-by-Event Variation}
\label{ch4:sec:individual event variation}

For clarity we have focused on the \textit{mean} evolution of individual enrichment events at fixed injection energy in the results thus far. We turn now to discuss how much variety exists among these individual events. For each of the discussed runs (from set \runonenu), Figure~\ref{ch4:fig:CGM_CNM_variance} shows the fraction of metals ejected from the galaxy (left) and the fraction of metals in the ISM within the CNM (center) as a function of source energy at two different times, 20~Myr (blue) and 100~Myr (orange). In the right panel, we show how the spread at each of these times varies for each event. The averaged values at each energy are shown as points, and the bars denote the min/max value for each event energy.\footnote{While the standard deviation or IQR may be more informative here, rather than min / max, the sample size at each energy, with the exception of the E46 runs, is too small to reliably do so.} As shown, there is substantial variation at each energy. This is particularly striking for the E46 events, which seemingly can have effectively any value for $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ at any point during the evolution. Finally, the right panel shows that the degree of inhomogeneous mixing for each event varies significantly as well. However, there is a more clear trend for each of the mean, min, and max values of the spread at 100~Myr (decreasing as a function of increasing $E_{\rm ej}$).

 \begin{figure*}
   \centering
   \includegraphics[width=0.975\linewidth]{figures/ch4/II_Eej_CNM_avg}
   \caption{A summary of the results presented in Figure~\ref{ch4:fig:CGM_CNM} and Figure~\ref{ch4:fig:CGM_CNM} that demonstrates the significant variation in $f_{\rm CGM}$, $f_{\rm CNM} / f_{\rm Disk}$, and the mean-median separate at fixed injection energy. The points give the average values at each energy at two different times, while the error bars show the min/max values at each energy. The points at different times have been offset slightly along the horizontal axes for clarity. Since they are only two points, we show the results from \runone E52\_r0 and \runone E52\_r300 without averaging.}
   \label{ch4:fig:CGM_CNM_variance}
 \end{figure*}

@@

\subsection{Dependence on Radial Position}
\label{ch4:sec:radial position}
We place each source in regular, but arbitrary, positions in the galaxy without consideration for the local ISM conditions at injection. Sources are placed at the center of the galaxy ($r = 0$~pc) or at various radii (at cardinal positions at 300~pc, and 600~pc), all in the mid-place of the galaxy. Since the E46 events were likely to not self-interact, we additionally placed events at $r = 100$~pc and one event at each $r$ at about one scale height above the disk ($z = 50$~pc). For each \runonenu~ run, we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ as a function of radial position in the galaxy for each event in Figure~\ref{ch4:fig:radial position}. The points, error bars, and colors are the same as in Figure~\ref{ch4:fig:CGM_CNM_variance}.

@@

We compute the average ISM properties within a 4-zone radius ($4 \times dx$=7.2~pc) around each injection site to examine any potential correlations in evolution with ISM properties.\footnote{Note that the injection region for each event is a 3-cell radius sphere mapped onto the grid with a CIC interpolation scheme.} In Figure~\ref{ch4:fig:ISM_variance} we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ for each \runonenu~ run as a function of the average number density ($<n>$) in the injection region within 0.1~Myr (our time resolution) of the event. The E46 events clearly do a better job sampling the ISM density PDF, with a slight trend of decreasing $f_{\rm CGM}$ (more retained metals) with local ISM number density. This difference is even true after 100~Myr of evoltuion, but still with significant variation at fixed $n$. The number statistics limit a conclusive result for the higher energy events, but generally there does not appear to be an obvious trend in behavior of $f_{\rm CGM}$ as a function of $n$. This is perhaps because these energies are sufficient to heat up and dramatically reduce the density of the ISM in the injection region anyway, such that $n$ before the event is not as important as it is for the E46 events which are much less dynamically important. $f_{\rm CNM} / f_{\rm Disk}$ does not display any significant variation with $n$ at 100~Myr, but is slightly elevated at 20~Myr at higher densities. There is still a variety of behavior at fixed $n$. Finally, we note no obvious trends of the spread of the metal PDFs as a function of $n$ for all but the E46 run. Again, the local ISM densities is likely not too dynamically relevant to the higher energy events -- hence their lack of variation with $n$ -- but this is not true for E46. There is a fairly clear trend of increasing inhomogeneity in the CNM as a function of the initial $n$ in the injection region for E46, that is not present at 20~Myr, but is obvious at 100~Myr.

@@

Finally, we examine the companion run (\runtwonu) to see if the global star formation rate of the galaxy drives variation in these results. Although this does not help to directly account for the variation seen in the previous plots, this could provide insight into how SFR (in an averaged sense) affects the evolution of metals. We note that correlation with SFR is of interest not because the formation of stars by themselves, but rather the increase in feedback associated with a higher SFR (and thus a warmer / hotter ISM, greater turbulence in the ISM, and more significant outflows). We compare these two runs in Figure~\ref{ch4:fig:SFR_comparison_CGM_CNM}, showing runs from \runonenu~ in green and runs at the lower SFR, \runtwonu, in purple. $f_{\rm CGM}$ increases much more rapidly across injection energies in \runonenu. However, this value seems to be similar across runs for E49 and E51 by $\sim$50 Myr, with an additional increase again in the \runonenu~ simulations towards the end of the time. This initial spike, lull, and second increase corresponds to the period of active star formation, slightly lull, and increase in SFR experienced during the \runonenu~ simulations (see Figure~\ref{ch1:figsfr_mass_evolution}; the lull occurs between 40 and 80 Myr after the enrichment events, or during 220 - 260 Myr in that figure). The SFR is consistently low (or zero) throughout \runtwonu, with just a couple brief increases in SFR. Although E49 and E51 exhibit similar behavior across global differences in SFR, the difference in E46 is significant throughout the examined time period. As these sources do not contain enough energy to eject their metals from the galaxy by themselves, they are only ejected from the galaxy by being swept up in the ISM during other feedback events. The evolution of these events are therefore much more dependent upon the global galaxy properties.

The fraction of metals contained within the CNM of the ISM does not seem to depend too strongly on the SFR, unlike $f_{\rm CGM}$. It is generally true that after the initial $\sim$50 Myr of enrichment, a greater fraction of the metals in the ISM are contained in the CNM during the run with the lower SFR (\runtwonu), but this difference is not large and is only significant for the higher energy events. In addition, the mass fraction of the CNM is greater during run \runtwonu~ than in run \runonenu, so it may simply be that the long-term evolution of the fraction of metals in the CNM is more dependent upon the phase structure of the ISM -- which is regulated by stellar feedback -- than the feedback directly.

\begin{figure*}
  \centering
  \includegraphics[width=0.45\linewidth]{figures/ch4/combined_CGM_average_evolution}
  \includegraphics[width=0.45\linewidth]{figures/ch4/combined_CNM_average_evolution}\\
  \includegraphics[width=0.45\linewidth]{figures/ch4/combined_CNM_average_mean-median}
  \caption{The same as Figure~\ref{ch4:fig:CGM_CNM}, but comparing across both sets of runs with different SFRs. Comparing across line colors for a fixed linestyle compares the difference the global SFR has on the average enrichment evolution for both $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$.}
  \label{ch4:fig:SFR_comparison_CGM_CNM}
\end{figure*}

\textbf{Discuss Mean-median evolution here once AGB400 is up}

@@

Galactic chemical evolution is far from ``one-zone" and metal mixing is clearly neither homogeneous nor instantaneous. As demonstrated in our simulations published as \cite{Emerick2018b} and as found in the analytic work of \cite{KrumholzTing2018}, the properties of metal mixing in the ISM of various galaxies depends upon the characteristics of the individual sources -- namely the energy with which they eject metals into the ISM. This sets the thermal phase in which these metals couple to most effectively, and the volume over which they are initial injected before additional mixing by global galactic dynamics begins to dominate. In short, lower energy enrichment events, like those from AGB winds, mix much more slowly throughout the ISM than metals from higher energy events, like supernovae. As found in \cite{Emerick2018b}, this difference is still present even when accounting for the fact that AGB events are both more common and evenly distributed in the galaxy than SNe. In the new work discussed in this chapter, we further explore this behavior to try and determine the evolution of individual enrichment sources -- rather than just the average behavior of many sources -- and to explore how $E_{\rm ej}$, radial position in the galaxy, local ISM density, and global galaxy SFR affects the evolution of these metals.
